\modLP{Identification of} Biological Regulatory Networks from Process Hitting models

### 2014-09-17 parts-tr2global.tex
sect 3.2

\todo{(LP) redo text}

Any sort that does not act as a component should then be treated as a cooperative sort.
As explained in \pref{ssec:PH}, the role of a cooperative sort $\upsilon$ is to compute the current
state of set of cooperating processes.
Hence, for each sub-state $\sigma$ formed by the sorts hitting $\upsilon$, $\upsilon$ should
converge to a focal process.
This is expressed by \pref{pro:wf-cooperative-sort}, where
the set of sorts having an action on a given sort $a$ is given by 
$\PHdirectpredec{a}$ (\pref{eq:ph_direct_predec})
and $\PHproc(\sigma)$ is the set of processes that compose the sub-state $\sigma$.

\begin{equation}
\forall a \in \PHs, \PHdirectpredec{a} \DEF \{b \in \PHs \mid \exists \PHfrappe{b_i}{a_j}{a_k}\in\PHa \}
\label{eq:ph_direct_predec}
\end{equation}

\begin{property}[Well-formed cooperative sort]\label{pro:wf-cooperative-sort}
A sort $\upsilon\in\PHs$ is a well-formed cooperative sort if and only if
each configuration $\sigma$ of its predecessors leads $\upsilon$ to a unique focal process,
denoted by $\upsilon(\sigma)$:
\[
\forall \sigma \in {\textstyle\prod_{
a\in\PHdirectpredec{\upsilon} \wedge a\neq \upsilon}}
\PHl_{a},
\focals(\upsilon,\PHl_\upsilon,\PHproc(\sigma)\cup \PHl_\upsilon) = \{ \upsilon(\sigma) \}\]
\end{property}

Such a property allows a large variety of definitions of a cooperative sort, but
for the sake of simplicity, does not allow the existence of multiple focal processes.
While this may be easily extended to (the condition becomes 
$\focals(\upsilon,\PHl_\upsilon, \PHproc(\sigma)\cup \PHl_\upsilon)\neq\emptyset$), it makes some
hereafter equations a bit more complex to read as they should handle a set of focal processes instead
of a unique focal process.

>>>>

As discussed at the beginning of this section, and as formally described in the above sub-section, 
in the scope of a given context,
a PH sort may converge to one among several different processes (indeterminism) or may never converge
(infinite loop).

\todo{this statement is too strong if we consider intervals, but correct if we are not}
It results that deriving Thomas specification for dynamics of a PH model
requires that every sort converges toward a unique focal process, i.e., that every sort has a
deterministic behaviour w.r.t. any context, as stated by \pref{pro:deterministic-sort}.

\begin{property}[Deterministic sort]\label{pro:deterministic-sort}
A sort $a\in\PHs$ is a deterministic sort if and only if
each configuration $\sigma$ of its predecessors leads $a$ to a unique focal process,
denoted $a(\sigma)$:
\[
\forall \sigma \in {\textstyle\prod_{
b\in\PHdirectpredec{\upsilon} \wedge b\neq a}}
\PHl_{b},
\focals(a,\PHl_a,\sigma) = \{ a(\sigma) \}\]
\end{property}

\todo{examples}
